Converting LaTeX Legacy R Journal Articles into R Markdown Articles using texor and rebib

### 2023-08-18 supplement/bibliography/rebib.tex
edits to bibliography supplement

% !TeX root = RJwrapper.tex
\title{Working with bibliography in texor}
\author{by Abhishek Ulayil}

@@

\section{Introduction}
The \CRANpkg{rebib} package is a spun-off from \CRANpkg{texor} package.

The decision to do this is the fact that the bibliography section 
in \CRANpkg{texor} package was expanding significantly, enough to deserve a dedicated space.

Thus \CRANpkg{rebib} was born and has improved over the past few months to include more advanced features and utilities.

@@

In LaTeX, we have multiple ways to declare bibliography entries, which include embedding bibliography entries in the document itself or isolating that section to an external file commonly called a bbl file. Then we have the BibTeX format, which has its structure and is much superior in managing bibliographic entries.

LaTeX has no problem dealing with multiple formats and works well with all these bibliographic formats. However, some web publishing journals like RJ web articles from RJournal do not support embedded bibliographies. As a result, they need a BibTeX equivalent of the same.

The \CRANpkg{rebib} package converts embedded LaTeX bibliographies into a close  BibTeX equivalent. This package eases the author from the manual conversion of the documents and aids the texor project in automating the conversion of past RJournal articles to the RJ-web-article format.

\subsection{features of rebib}

\begin{itemize}
  \item Reads bib chunks to produce a very close BibTeX equivalent
  \item Title and author are usually mandatory fields
  \item URL, ISBN, publisher, pages and year are optional fields and will be enabled when relevant
  \item Rest of the data is stored in \verb|"journal"|(internally) and \verb|"publisher"|(when writing BibTeX file)
  \item Ignores commented LaTeX code 
  \item Citation tracker
  \item Bibliography Aggregation
\end{itemize}

@@

The magic behind \CRANpkg{rebib} is a parser which parses bibliographic data and assorts them according to the matching regular expressions.

@@

This stage is a minor step where it reads the Embedded Bibliography from the LaTeX document. This step also includes Filtering out the commented code to avoid un-intended entries read.

Lastly, the data is broken down based on the LaTeX macro \verb|\\bibitem| as a marker for a new entry and this assorted data is exported to a variable.

@@

The first value to be parsed is the \verb|unique_id| also called the citation reference which is used to cite elements inside the article. Usually, this is in the first or second line of the whole entry. The position of the 
\verb|unique_id| will determine the position of the author names.

@@

This way the crucial elements of the bibliographic entry (unique\_id, author names and title ) are parsed out.
The remaining data is stored as \verb|journal| internally and \verb|publisher| when writing to a new BibTeX file.

@@

A series of filters for ISBN, URL, pages and year fields are applied to search for relevant data from the remaining data. If the data is not available then it is set as NULL and will be skipped while writing the BibTeX file. 
There is a lot of filtering of common LaTeX elements and after that, the data remaining is stored in a structured format to be written to a file.

@@

I expect the authors who are converting the document to take a look and check if there are any errors or missing values.

\section{General usage}
In general to use \CRANpkg{rebib} package, it is quite easy with requiring only one function call.
\subsubsection{Using with RJournal article}
In bibliography construction mode, \CRANpkg{rebib} will check for existing BibTeX files and will convert the embedded bibliography only when there are no linked BibTeX files in the RJ article.
\begin{verbatim}
%% typically rebib will search for this line in the RJ article, 
%% not the RJwrapper file
\bibliography{example}
\end{verbatim}
If there is no linked BibTeX file, \CRANpkg{rebib} will begin its procedure of converting the embedded bibliography to BibTeX and then link it with the article file as well.
\begin{Schunk}
\begin{Sinput}
# for files without BibTeX source
rebib::handle_bibliography(your_article_path)
#> Warning in get_bib_file(article_dir, file_name): No Bib files found !
#> BibTeX file does not exist
#> will parse for bibliography
#> bibtex file created
cat(readLines(paste(your_article_path,"example.bib",sep="/")),sep = "\n")
#> @book{ihaka:1996,
#> author = {{Ihaka, Ross and Gentleman, Robert}},
#> title = {{R: A Language for Data Analysis and Graphics}},
#> publisher = {Journal of Computational and Graphical Statistics 3:   :},
#> pages = {299--314},
#> year = {1996},
#> url = {https://doi.org/10.1080/10618600.1996.10474713}
#> }
#> @book{R,
#> author = {R {Core Team}},
#> title = {{R: A Language and Environment for Statistical Computing}},
#> publisher = {R Foundation for Statistical Computing Vienna Austria    :},
#> year = {2016},
#> url = {https://www.R-project.org/},
#> isbn = {3-900051-07-0}
#> }
\end{Sinput}
\end{Schunk}

@@

\subsubsection{Using with any LaTeX/BBL file}
The bibliography converter will convert any LaTex or BBL file given as a path. This way, one can generate BibTeX from the embedded bibliography or BBL file.
\begin{verbatim}
rebib::biblio_converter(path_to_your_file)
#> working directory : /temp/Rtmpt6Qk0v/exampledir/standalone
#> bib entries : 6
#> Written BibTeX file : /temp/Rtmpt6Qk0v/exampledir/standalone/sample.bib
\end{verbatim}

@@

In summary the \CRANpkg{rebib} package supports:
\begin{itemize}
\item Conversion of embedded natbib type bibliography used in RJournal
\item Aggregation of bibliographic entries from the article as well as BibTeX
\end{itemize}

>>>>

% !TeX root = RJwrapper.tex
\title{Handling bibliographies with texor}
\author{by Abhishek Ulayil}

@@

\section{Introduction}
The \CRANpkg{rebib} package is a spin-off from the \CRANpkg{texor} package.

The spin-off cam about because the bibliography section 
in \CRANpkg{texor} package was expanding significantly, enough to deserve a dedicated space.

Thus \CRANpkg{rebib} was born and it has improved over the past few months to include more advanced features and utilities.

@@

In LaTeX, we have multiple ways to declare bibliography entries, which include embedding entries formatted to a given style in the document itself or isolating that section to an external file commonly called a \file{.bbl} file. Then we have the BibTeX format, which is designed to store the full bibliographic information for each reference, making it superior for managing references.

LaTeX has no problem dealing with multiple formats and works well with all these bibliographic formats. However, some web publishing journals like R Markdown articles for the R Journal do not support embedded bibliographies. As a result, they need a BibTeX equivalent of the same.

The \CRANpkg{rebib} package converts embedded LaTeX bibliographies into a close  BibTeX equivalent. This package relieves the author from manual conversion of the documents and aids the project to convert legacy R Journal articles to the current R Markdown format.

\subsection{Features of rebib}

\begin{itemize}
  \item Extracts embedded bibliographic entries from a \file{.tex} file.
  \item Creates the mandatory title and author fields.
  \item Creates the optional URL, ISBN, publisher, pages and year fields, when available.
  \item Stores the remaining information in \verb|"journal"| (internally) and \verb|"publisher"| (when writing BibTeX file).
  \item Ignores commented LaTeX code. 
  \item Tracks citations included in the document.
  \item Aggregates embedded bibliographic entries with references from supplementary \file{.bbl} and \file{.bib} files.
\end{itemize}

@@

The magic behind \CRANpkg{rebib} is a parser which parses bibliographic data and sorts them according to the matching regular expressions.

@@

This stage is a minor step where it reads the Embedded Bibliography from the LaTeX document. This step also includes filtering out commented lines to avoid reading un-intended entries.

The bibliography is split into entries using the LaTeX macro \verb|\\bibitem| as a marker for a new entry and the resulting data is exported to a variable.

@@

The first value to be parsed is the \verb|unique_id|, also called the citation reference which is used to cite elements inside the article. Usually, this is in the first or second line of the whole entry. The position of the 
\verb|unique_id| will determine the position of the author names.

@@

This way the crucial elements of the bibliographic entry (unique\_id, author names and title) are parsed out.
The remaining data is stored as \verb|journal| internally and \verb|publisher| when writing to a new BibTeX file.

@@

A series of filters for ISBN, URL, pages and year fields are applied to search for relevant data from the remaining data. If the data is not available, then it is set as \code{NULL} and will be skipped while writing the BibTeX file. 
There is a lot of filtering of common LaTeX elements and after that, the remaining data is stored in a structured format to be written to a file.

@@

Authors who are converting a bibliography from \file{bbl} should check if there are any errors or missing values.

\section{General usage}

General use of the \CRANpkg{rebib} package is quite easy, requiring only one function call.

\subsubsection{Using with RJournal article}

In bibliography construction mode, \CRANpkg{rebib} will check for existing BibTeX files and will convert the embedded bibliography only when there are no linked BibTeX files in the RJ article.
\begin{verbatim}
%% typically rebib will search for this line in the RJ article, 
%% not the RJwrapper file
\bibliography{example}
\end{verbatim}
If there is no linked BibTeX file, \CRANpkg{rebib} will begin its procedure of converting the embedded bibliography to BibTeX and then link it with the article file as well.
\begin{Schunk}
\begin{Sinput}
# for files without BibTeX source
rebib::handle_bibliography(your_article_path)
#> Warning in get_bib_file(article_dir, file_name): No Bib files found !
#> BibTeX file does not exist
#> will parse for bibliography
#> bibtex file created
cat(readLines(paste(your_article_path,"example.bib",sep="/")),sep = "\n")
#> @book{ihaka:1996,
#> author = {{Ihaka, Ross and Gentleman, Robert}},
#> title = {{R: A Language for Data Analysis and Graphics}},
#> publisher = {Journal of Computational and Graphical Statistics 3:   :},
#> pages = {299--314},
#> year = {1996},
#> url = {https://doi.org/10.1080/10618600.1996.10474713}
#> }
#> @book{R,
#> author = {R {Core Team}},
#> title = {{R: A Language and Environment for Statistical Computing}},
#> publisher = {R Foundation for Statistical Computing Vienna Austria    :},
#> year = {2016},
#> url = {https://www.R-project.org/},
#> isbn = {3-900051-07-0}
#> }
\end{Sinput}
\end{Schunk}

@@

\subsubsection{Using with any LaTeX/.bbl file}

The bibliography converter will convert any LaTeX or \file{.bbl} file given as a path. This way, one can generate BibTeX from the embedded bibliography or \file{.bbl} file.
\begin{verbatim}
rebib::biblio_converter(path_to_your_file)
#> working directory : /temp/Rtmpt6Qk0v/exampledir/standalone
#> bib entries : 6
#> Written BibTeX file : /temp/Rtmpt6Qk0v/exampledir/standalone/sample.bib
\end{verbatim}

@@

In summary the \CRANpkg{rebib} package supports:
\begin{itemize}
\item Conversion of embedded natbib type bibliography entries used in the R Journal
\item Aggregation of bibliographic entries from the article as well as separate BibTeX files
\end{itemize}

### 2023-08-18 texor.Rmd
edits to bibliography supplement

- extracting `bbl` chunks from a `.tex` file;
- creating the mandatory title and author fields;
- creating the optional URL, ISBN, publisher, pages and year fields, when available;
- storing remaining information in `"journal"` (internally) and `"publisher"` (when writing BibTeX file);
- ignoring commented LaTeX code;
- tracking citations included in the document, and
- aggregating references from embedded `bbl` with references from supplementary `.bbl` and `.bib` files.

>>>>

- extracting embedded bibliographic entries from a `.tex` file;
- creating the mandatory title and author fields;
- creating the optional URL, ISBN, publisher, pages and year fields, when available;
- storing remaining information in `"journal"` (internally) and `"publisher"` (when writing BibTeX file);
- ignoring commented LaTeX code;
- tracking citations included in the document, and
- aggregating references from embedded bibliographic entries with references from supplementary `.bbl` and `.bib` files.
